---
title: "Merging the datasets with tidyverse"
author: ""
date: "07/02/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Day 2

We are going to use the covid 19 dataset along with the other two datasets (in "data/" folder). 

- Vaccination data
- Country level demographic data
- Presidential voting in 2020
 (or you can use, Presidential voting in 2016)

These datasets are coming from the following places

- https://data.cdc.gov/Vaccinations/COVID-19-Vaccinations-in-the-United-States-County/8xkx-amqh
- https://www.kaggle.com/stevepalley/2016uspresidentialvotebycounty
- https://www.kaggle.com/muonneutrino/us-census-demographic-data
- https://www.kaggle.com/unanimad/us-election-2020
  - (for this one, I added fips to the original file and remove quite a bit of rows)


The task is

1. check the data completeness (e.g. make sure that all datasets have common list of counties)
2. merge the datasets
3. check whether there are any relations between covid and any demographic/political indexes


#### Download the file
The vaccination data is here: https://www.dropbox.com/s/3bqwbxw6mr9t9jv/COVID-19_Vaccinations_in_the_United_States_County.csv.tar.gz?dl=1

```{r}
dir.create("tmp")
download.file("https://www.dropbox.com/s/3bqwbxw6mr9t9jv/COVID-19_Vaccinations_in_the_United_States_County.csv.tar.gz?dl=1", destfile = "tmp/COVID-19_Vaccinations_in_the_United_States_County.csv.tar.gz")

```

## Packages

```{r}
library(tidyverse)
```

## Read three data files

```{r}
df_vac <- vroom::vroom("tmp/COVID-19_Vaccinations_in_the_United_States_County.csv.tar.gz")
df_demo <- read_csv("data/acs2017_county_data.csv")
df_pres <- read_csv("data/pres20results.csv")
```


## Check the validity of the data

### How the data look like


```{r}
library(magrittr)
df_vac %>% head()
df_vac %>% dim()

#df_vac %<>%
#df_vac <- df_vac %>%
  
df_vac %<>%
  mutate(Date = Date %>% as.Date(format = "%m/%d/%Y"))
df_vac %<>%
  mutate(FIPS = FIPS %>% as.integer) %>%
  filter(!is.na(FIPS))



```


```{r}
df_demo %>% head()
df_demo %<>%
  rename(FIPS = CountyId)
df_demo 
```


```{r}
df_pres %>% head()
df_pres %<>% 
  rename(FIPS = fips)

```
- renaming is necessary for `data_demo`
- calculating voting percentage from the presidential election data


```{r}

df_pres %<>%
  filter(candidate == "Donald Trump") %>%
  mutate(vote_pct  = votes/total_votes)
df_pres
```



### How many rows in each df?

```{r}

```

### Number of counties in each data frame

- Which variable would you use to check?
- How many unique counties in each data set?
- Do they match? (if not does this seem problematic)?
    
```{r}
df_vac_counties <- df_vac %>% count(FIPS, Recip_County, Recip_State)

df_vac_counties %>%
  anti_join(df_demo, by = 'FIPS')


df_vac_counties %>%
  anti_join(df_pres, by = 'FIPS')

```

## Keep the necessary data

### Get the latest numbers cases

- From covid dataset, get the latest figures for each county

```{r}

df_vac_latest <- df_vac %>%
  group_by(FIPS) %>%
  filter(Date == max(Date)) %>%
  ungroup() 

```


### Extract trump votes

presidential votes
  - Keep only trump vote rows

```{r}
# already done
```

## Merge the datasets


### Merging

- Now lets merge the data set
  1. latest vaccination data
  2. country demographics
  3. presidential votes
  
```{r}
df_merged <- df_vac_latest %>%
  inner_join(df_pres, by = "FIPS") %>%
  inner_join(df_demo, by = "FIPS") 

```

## Using the merged datasets 

- Get a correlation matrix or plot for the data
- `cor()`
- `pairs()`
- `GGally::ggpairs()`, correlation plot (I showed in day 4 lecture)

```{r}
df_merged %>%
  select(Series_Complete_Pop_Pct, Booster_Doses_18Plus_Vax_Pct,
         vote_pct, Hispanic, Black, IncomePerCap, Unemployment) %>% 
  cor(use = "complete.obs")
df_merged %>%
  select(Series_Complete_Pop_Pct,
         vote_pct, Hispanic, Black, IncomePerCap, Unemployment) %>% 
  pairs()

```
## Month-by-month correlations between Trump Vote and caccination takeup.

- Let's calculate the monthly update of correlation between county level vaccination and Trump vote.
- How to address it?
  1. Filter first date of month vaccination data
  2. Merge with presidential vote
  3. Group data, calculate correnation each month
  4. Plot?


```{r}
df_vac %>%
  mutate(Date = as.Date(Date)) %>%
  filter(lubridate::day(Date) == 1) %>%
  inner_join(df_pres, by = "FIPS") %>%
  group_by(Date) %>%
  summarize(cor_vote_vac = cor(vote_pct, Series_Complete_Pop_Pct, use = "pairwise.complete.obs")) %>%
  ggplot(aes(x = Date, y = cor_vote_vac)) + geom_line()

```

## Create a database

From the three datasets, let's create a database

What we need to do is:

1. Remove some duplicated fiels
2. Create a database connection using `DBI` and `RSQlite`
3. Create tables from these datasets (using `dbWriteTable()`)
4. Disconnect from the db

### check the database fields

- Remove duplicate rows
- Convert dates to character

```{r}
df_vac %>% distinct() %>% dim()
```


### Create database connection

```{r}
library(DBI)
file.remove("~/covid.sqlite")
db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")

```

### Create/populate tables

```{r}
df_demo <- df_demo %>% select(-c(State, County))
df_pres <- df_pres %>% select(-c(party, state, candidate, county))
df_vac <- df_vac %>% mutate(Date = as.character(Date))

dbWriteTable(db, name = "demo", df_demo)
dbWriteTable(db, name = "trump", df_pres)
dbWriteTable(db, name = "vac", df_vac)

```

### Disconnect

```{r}
dbDisconnect(db)
```
