---
title: "Using database"
author: ""
date: "07/02/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Connect to the database

- Now let's use the databases
- First connect to the database

```{r}
library(DBI)
library(tidyverse)

db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")
```

## Run queries

### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
dbGetQuery(db, "SELECT * FROM sqlite_master WHERE type='table';")
```


```{r}
dbGetQuery(db, "SELECT * FROM demo LIMIT 5")
```


```{r}
dbGetQuery(db, "SELECT * FROM vac LIMIT 5")
```
```{r}
dbGetQuery(db, "SELECT * FROM trump LIMIT 5")
```

### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2021, and select county with less than 10 pct vaccination

```{r}
dbGetQuery(db, "SELECT Date, COUNT() FROM vac GROUP BY Date") %>%
  arrange(desc(Date))


dbGetQuery(db, "SELECT * FROM vac WHERE Date = '2022-07-27'") 
dbGetQuery(db, "SELECT * FROM vac WHERE Date = '2021-08-01' AND Series_Complete_Pop_Pct > 10") 

```
#### Ordering and aggregation

1. Order counties with the highest rate of vaccination on Aug 14, 2021
2. Calculate the state level vaccine count


```{r}
dbGetQuery(db, "SELECT * FROM vac WHERE Date = '2021-08-14' ORDER BY Series_Complete_Pop_Pct DESC") 
dbGetQuery(db, "SELECT Recip_State, SUM(Series_Complete_Yes) FROM vac WHERE Date = '2021-08-14' GROUP BY Recip_State") 

```


### Merging tables

- merge vaccine and demo table by fips

```{r}
dbGetQuery(db, "SELECT vac.*, demo.TotalPop FROM vac 
            INNER JOIN demo ON vac.FIPS = demo.FIPS WHERE Date = '2021-08-14' ")

```

## dbplyr

Now we will use `dbplyr`

### Get the object for each table

```{r}
tb_demo <- tbl(db, "demo")
tb_trump <- tbl(db, "trump")
tb_vac <- tbl(db, "vac")
  
```


### Let's check the same thing we did:


#### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
tb_demo %>% head(10) #%>% show_query()
```


#### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 14 Aug, 2021, and select county with more than 10 percent vaccination rate

```{r}

tb_vac %>% 
  count(Date) %>%
  arrange(desc(Date))


tb_vac %>%
  filter(Date == "2022-07-27") #%>% show_query()


tb_vac %>%
  filter(Date == "2021-08-14") %>% 
  filter(Series_Complete_Pop_Pct > 10) #

```
#### Ordering and aggregation

1. Order counties with the highest number of vaccine percentage
2. Calculate the state level vaccination total


```{r}
dbGetQuery(db, "SELECT * FROM vac WHERE Date = '2021-08-14' ORDER BY Series_Complete_Pop_Pct DESC") 
dbGetQuery(db, "SELECT Recip_State, SUM(Series_Complete_Yes) FROM vac WHERE Date = '2021-08-14' GROUP BY Recip_State") 

tb_vac  %>%
  filter(Date == "2021-08-14") %>% 
  arrange(desc(Series_Complete_Pop_Pct)) %>%
  head()

tb_vac %>%
  filter(Date == "2021-08-14") %>%
  group_by(Recip_State) %>%
  summarise(sum(Series_Complete_Yes)) #%>% show_query()
```


### Merging tables

- merge vaccine and demo table by fips
- calculate vaccination rate manually from Series_Complete_Yes and TotalPop
  - compare with Series_Complete_Pop_Pct


```{r}

tb_vac %>%
  inner_join(tb_demo, by = "FIPS")

```


```{r}
tmp <- tb_vac %>%
  filter(Date == "2021-08-14") %>%
  group_by(Recip_State) %>%
  summarise(sum(Series_Complete_Yes)) #%>% show_query()
```


```{r}
tmp %>% class()
```

```{r}
df_tmp <- tb_vac %>%
  filter(Date == "2021-08-14") %>%
  group_by(Recip_State) %>%
  summarise(sum(Series_Complete_Yes)) %>% as_tibble()

df_tmp2 <- tb_vac %>%
  filter(Date == "2021-08-14") %>%
  group_by(Recip_State) %>%
  summarise(sum(Series_Complete_Yes)) %>% collect()

df_tmp2 

```


